Timely Object Recognition

### 2012-11-03 writeups/nips12/evaluation.tex
camera-ready draft sent to collaborators

We evaluate our system on the multi-class, multi-label detection task as previously described.
Each detection episode takes an image and outputs detections with associated times, based on the order of actions taken.
We evaluate on a popular detection challenge task: the PASCAL VOC 2007 dataset \cite{pascal-voc-2010}, which exhibits only a rather modest amount of class co-occurrence: the ``person'' class is highly likely to occur, and less than $10\%$ of the images have more than two classes.

@@

For the detector actions, we use one-vs-all cascaded deformable part-model detectors on a HOG featurization of the image \cite{Felzenszwalb2010b}, with associated linear classification of the detections.
There are $20$ classes in the PASCAL challenge task, so there are $20$ detector actions.
Running a detector on a PASCAL image takes about $1$ second.

We test three different settings of the start and deadline times.
In the first one, the start time is immediate and execution is cut off at $20$ seconds, which is enough time to run basically all actions.
In the second one, execution is cut off after only $10$ seconds.
Lastly, we measure performance between $5$ seconds and $15$ seconds.
These operating points show how our method behaves when deployed in different conditions.
The results are given in rows of \autoref{tab:results}.

\begin{figure}[h!]
\centering
\subfloat[]{
\includegraphics[width=0.47\linewidth]
    {../figures/final1_515.png}\label{fig:results1}} \hfill
\subfloat[]{
\includegraphics[width=0.47\linewidth]
    {../figures/apvst_expl.pdf}\label{fig:rewards}}
\caption{(a) AP vs. Time curves for Random, Oracle, the Fixed Order baseline, and our best-performing policy, with start time of 5 seconds and deadline of 15 seconds. (b) Reward function explanation.}
\end{figure}

We establish the first baseline for our system by selecting actions randomly at each step.
As shown in Figure \autoref{fig:results1}, this \textbf{Random} policy results in a roughly linear gain of AP vs. time.
This is expected: the detectors are capable of obtaining a certain level of performance; if half the detectors are run, the expected performance level is half of the maximum level.

To establish an upper bound on performance, we plot the \textbf{Oracle} policy, obtained by re-ordering the actions with the hindsight of the results they produced.

@@

Then there are our proposed systems, described in the previous section : \textbf{RL w/ Direct} and \textbf{RL w/ MRF}.
In experiments, the two were not exceedingly different in performance, with the MRF model consistently performing slightly better.

@@

Lastly, we include an additional scene-level GIST feature that updates the posterior probabilities of all classes.
This is considered one action, and takes about $0.3$ seconds.
This setting always uses the MRF model to properly update the class probabilities with GIST observations.
This brings another boost in performance.
All results are shown in \autoref{tab:results}.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.86\linewidth]{../figures/trajectories.pdf}
  \caption{
Visualizing the action trajectories of different policies.
Action selection traces are plotted in orange over many episodes; the size of the blue circles correspond to the $\Delta ap$ obtained by the action.
We see that the \textbf{Random} policy selects actions and obtains rewards randomly, while the \textbf{Oracle} policy obtains all rewards in the first few actions.
The \textbf{Fixed Order} policy selects actions in a static optimal order.
Our policy does not stick a static order but selects actions dynamically to maximize the rewards obtained early on.
}
  \label{fig:trajectories}
\end{figure}

Additionally, we evaluated different settings of the discount parameter $\gamma$, and found that a mid-range value works best.
We only report results for that value on the test set.
\autoref{fig:weights} shows how changing $\gamma$ affects the learned weights: the GIST action is learned to never be taken in the greedy setting, because it does not bring any immediate detections results.
In the reinforcement learning setting, however (higher $\gamma$), GIST is learned to be taken for its value in quickly refining the class presence probabilities.

\autoref{fig:trajectories} illustrates the action trajectories of different policies.

\begin{table}[t]
\caption{
The areas under the AP vs. Time curve for different experimental conditions.
Our method reliably beats the intelligent static baseline; including the GIST action further boosts performance.}
\label{tab:results}
\begin{center}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
Bounds & Random & Fixed Order & RL w/ MRF & w/ GIST         & Oracle \\ \hline
(0,20) & 0.250  & 0.342       & 0.378     & \textbf{0.382}  & 0.488 \\ 
(0,10) & 0.119  & 0.240       & 0.266     & \textbf{0.267}  & 0.464 \\ 
(5,15) & 0.257  & 0.362       & 0.418     & \textbf{0.420}  & 0.530 \\ \hline
\end{tabular}
\end{center}
\end{table}

>>>>

We evaluate our system on the multi-class, multi-label detection task, as previously described.
%Each detection episode takes an image and outputs detections with associated times, based on the order of actions taken.
We evaluate on a popular detection challenge task: the PASCAL VOC 2007 dataset \cite{pascal-voc-2010}.
This datasets exhibits a rather modest amount of class co-occurrence: the ``person'' class is highly likely to occur, and less than $10\%$ of the images have more than two classes.

@@

For the detector actions, we use one-vs-all cascaded deformable part-model detectors on a HOG featurization of the image \cite{Felzenszwalb2010b}, with linear classification of the list of detections as described in the previous section.
There are $20$ classes in the PASCAL challenge task, so there are $20$ detector actions.
Running a detector on a PASCAL image takes about $1$ second.

We test three different settings of the start and deadline times.
In the first one, the start time is immediate and execution is cut off at $20$ seconds, which is enough time to run all actions.
In the second one, execution is cut off after only $10$ seconds.
Lastly, we measure performance between $5$ seconds and $15$ seconds.
These operating points show how our method behaves when deployed in different conditions.
The results are given in rows of \autoref{tab:results}.

\begin{figure}[h!]
\centering
\subfloat[]{
\includegraphics[width=0.53\linewidth]
    {../figures/final1_515.png}\label{fig:results1}} \hfill
\subfloat[]{
\includegraphics[width=0.44\linewidth]
    {../figures/apvst_expl.pdf}\label{fig:rewards}}
\caption{
(a) AP vs. Time curves for Random, Oracle, the Fixed Order baseline, and our best-performing policy. (b) Graphically representing our reward function, as described in~\autoref{sec:value}.
}
\end{figure}

We establish the first baseline for our system by selecting actions randomly at each step.
As shown in Figure~\autoref{fig:results1}, the \textbf{Random} policy results in a roughly linear gain of AP vs. time.
This is expected: the detectors are capable of obtaining a certain level of performance; if half the detectors are run, the expected performance level is half of the maximum level.

To establish an upper bound on performance, we plot the \textbf{Oracle} policy, obtained by re-ordering the actions at the end of each detection episode in the order of AP gains they produced.

@@

Then there are instantiations of our method, as described in the previous section : \textbf{RL w/ Direct} inference and \textbf{RL w/ MRF} inference.
In experiments, the two were not exceedingly different in performance, with the \textbf{MRF} model consistently outperforming \textbf{Direct}.

@@

Lastly, we include an additional scene-level GIST feature that updates the posterior probabilities of all classes.
This is considered one action, and takes about $0.3$ seconds.
This setting always uses the MRF model to properly update the class probabilities with GIST observations.
This brings another small boost in performance.
The results are shown in \autoref{tab:results}.

Visualizing the learned weights in~\autoref{fig:weights}, we note that the GIST action is learned to never be taken in the greedy ($\gamma=0$) setting, but is learned to be taken with a higher value of $\gamma$.
It is additionally informative to consider the action trajectories of different policies in~\autoref{fig:trajectories}.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.9\linewidth]{../figures/trajectories.pdf}
  \caption{
Visualizing the action trajectories of different policies.
Action selection traces are plotted in orange over many episodes; the size of the blue circles correspond to the $\Delta ap$ obtained by the action.
We see that the \textbf{Random} policy selects actions and obtains rewards randomly, while the \textbf{Oracle} policy obtains all rewards in the first few actions.
The \textbf{Fixed Order} policy selects actions in a static optimal order.
Our policy does not stick a static order but selects actions dynamically to maximize the rewards obtained early on.
}
  \label{fig:trajectories}
\end{figure}

\begin{table}[t]
\caption{The areas under the AP vs. Time curve for different experimental conditions.}
\label{tab:results}
\begin{center}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
Bounds & Random & Fixed Order & RL w/ MRF & w/ GIST         & Oracle \\ \hline
(0,20) & 0.250  & 0.342       & 0.378     & \textbf{0.382}  & 0.488 \\ 
(0,10) & 0.119  & 0.240       & 0.266     & \textbf{0.267}  & 0.464 \\ 
(5,15) & 0.257  & 0.362       & 0.418     & \textbf{0.420}  & 0.530 \\ \hline
\end{tabular}
\end{center}
\end{table}

### 2012-11-03 writeups/nips12/introduction.tex
went through related work
camera-ready draft sent to collaborators

\section{Introduction}

@@

\begin{figure}[ht!]
\center{\includegraphics[width=0.86\linewidth]
    {../figures/figure1.pdf}}
  \caption{
A sample trace of our method.
At each time step beginning at $t=0$, potential actions are considered according to their \emph{value}, and the $\arg\max$ action is picked.

The selected action is performed and returns \emph{observations}.
Different actions return different observations: a detector returns a list of detections, while a scene context action simply returns its cmoputed feature. 
The \emph{belief model} of our system is updated with the observations, which influences the selection of the next action.

The final evaluation of a detection episode is the area of the \emph{AP vs. Time curve} between given start and end times.
The value of an action is the expected result of final evaluation if the action is taken and the policy continues to be followed, which allows actions without an immediate benefit to be scheduled.
}
  \label{fig:figure1}
\end{figure}

% TODO: more fully explain that this is a detection system
A hypothetical system for vision-based advertising presents a case study: companies pay money to have their products detected in images on the internet.
The system has different values (e.g. in terms of cost-per-click) and accuracies for different classes of objects, and the queue of unprocessed images varies in size.
The detection strategy to maximize profit in such an environment has to exploit every inter-object context signal available to it, because there is not enough time to run detection for all classes.

@@

Taking the task of object detection, we propose a new \emph{timeliness} measure of performance vs. time (shown in Figure~\ref{fig:figure1}).
We present a method that treats different detectors and classifiers as black boxes, and uses reinforcement learning to learn a dynamic policy for selecting actions to achieve the highest performance under this evaluation.
Evaluating on the PASCAL VOC dataset and evaluation regime, we are able to obtain better performance than all baselines when there is less time available than is needed to exhaustively run all detectors.

>>>>

\section{Introduction} \label{sec:introduction}

@@

\begin{figure}[ht!]
\center{\includegraphics[width=0.86\linewidth]
    {../figures/figure1.pdf}}
  \caption{
A sample trace of our method.
At each time step beginning at $t=0$, potential actions are considered according to their \emph{value}, and the $\arg\max$ action is picked.
The selected action is performed and returns \emph{observations}.
Different actions return different observations: a detector returns a list of detections, while a scene context action simply returns its cmoputed feature. 
The \emph{belief model} of our system is updated with the observations, which influences the selection of the next action.
The final evaluation of a detection episode is the area of the \emph{AP vs. Time curve} between given start and end times.
The value of an action is the expected result of final evaluation if the action is taken and the policy continues to be followed, which allows actions without an immediate benefit to be scheduled.
}
  \label{fig:figure1}
\end{figure}

A hypothetical system for vision-based advertising presents a case study: companies pay money to have their products detected in images on the internet.
The system has different values (in terms of cost per click) and accuracies for different classes of objects, and the queue of unprocessed images varies in size.
The detection strategy to maximize profit in such an environment has to exploit every inter-object context signal available to it, because there is not enough time to run detection for all classes.

@@

Taking the task of object detection, we propose a new \emph{timeliness} measure of performance vs. time (shown in Figure~\ref{fig:figure1}).
We present a method that treats different detectors and classifiers as black boxes, and uses reinforcement learning to learn a dynamic policy for selecting actions to achieve the highest performance under this evaluation.

Specifically, we run scene context and object class detectors over the whole image sequentially, using the results of detection obtained so far to select the next actions.
Evaluating on the PASCAL VOC dataset and evaluation regime, we are able to obtain better performance than all baselines when there is less time available than is needed to exhaustively run all detectors.

### 2012-11-03 writeups/nips12/nips2012.tex
working on nips final and reorganizing slides into deadline folders
camera-ready draft sent to collaborators

\documentclass{article} % For LaTeX2e
\usepackage{nips12submit_e,times}
\usepackage{graphicx}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{color}
\usepackage{subfig}
\usepackage{natbib}
\usepackage[pagebackref=true,breaklinks=true,colorlinks,bookmarks=false]{hyperref}

@@

\author{
Sergey Karayev \And
Tobias Baumgartner \And
Mario Fritz \And
Trevor Darrell
}

@@

%\nipsfinalcopy % Uncomment for camera-ready version

@@

\begin{abstract}
In a large visual multi-class detection framework, the timeliness of results can be crucial.
Our method for \emph{timely} multi-class detection aims to give the best possible performance at any single point after a start time; it is terminated at a deadline time.
Toward this goal, we formulate a dynamic, \emph{closed-loop} policy that infers the contents of the image in order to decide which detector to deploy next.
In contrast to previous work, our method significantly diverges from the predominant greedy strategies, and is able to learn to take actions with deferred values.
We evaluate our method with a novel \emph{timeliness} measure, computed as the area under an Average Precision vs. Time curve.
Experiments are conducted on the eminent PASCAL VOC object detection dataset.
If execution is stopped when only half the detectors have been run, our method obtains $66\%$ better AP than a random ordering, and $14\%$ better performance than an intelligent baseline.
On the timeliness measure, our method obtains at least $11\%$ better performance.
Our code, to be made available upon publication, is easily extensible as it treats detectors and classifiers as black boxes and learns from execution traces using reinforcement learning.
\end{abstract}

@@

\section{Conclusion}
We presented a method for learning ``closed-loop'' policies for multi-class object recognition, given existing object detectors and classifiers and a metric to optimize.
The method learns the optimal policy using reinforcement learning, by observing execution traces in training.
If detection on an image is cut off after only half the detectors have been run, our method does $66\%$ better than random, and $14\%$ better than an intelligent baseline. In particular, our method learns to take action with no intermediate reward in order to improve the overall performance of the system.

As always with reinforcement learning problems, defining the reward function requires some manual work.
Here, we derive it for the novel detection AP vs. Time evaluation that we suggest is most useful for evaluating efficiency in recognition.
Although computation devoted to scheduling actions is much less significant than the computation due to running the actions, an interesting future research direction is to explicitly consider this decision-making cost.
At time of publication, we will release the code of our modular framework, that allows for easy incorporation of other classifiers and detectors on different tasks and rewards.

% \subsubsection*{Acknowledgments}
% Use unnumbered third level headings for the acknowledgments. All
% acknowledgments go at the end of the paper. Do not include 
% acknowledgments in the anonymized submission, only in the 
% final paper.

>>>>

\documentclass{article} % For LaTeX2e
\usepackage{nips12submit_e,times}
\usepackage{graphicx}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{color}
\usepackage{subfig}
\usepackage{natbib}
\usepackage{array}
\usepackage{tabularx}
\usepackage[pagebackref=true,breaklinks=true,colorlinks,bookmarks=false]{hyperref}

@@

\author{
Sergey Karayev \\
UC Berkeley
\And
Tobias Baumgartner \\
Affiliation
\And
Mario Fritz \\
MPI Saarbr\"{u}cken
\And
Trevor Darrell \\
UC Berkeley
}

@@

\nipsfinalcopy % Uncomment for camera-ready version

@@

\begin{abstract}
In a large visual multi-class detection framework, the timeliness of results can be crucial.
Our method for \emph{timely} multi-class detection aims to give the best possible performance at any single point after a start time; it is terminated at a deadline time.
Toward this goal, we formulate a dynamic, \emph{closed-loop} policy that infers the contents of the image in order to decide which detector to deploy next.
In contrast to previous work, our method significantly diverges from the predominant greedy strategies, and is able to learn to take actions with deferred values.
We evaluate our method with a novel \emph{timeliness} measure, computed as the area under an Average Precision vs. Time curve.
Experiments are conducted on the eminent PASCAL VOC object detection dataset.
If execution is stopped when only half the detectors have been run, our method obtains $66\%$ better AP than a random ordering, and $14\%$ better performance than an intelligent baseline.
On the timeliness measure, our method obtains at least $11\%$ better performance.
Our method is easily extensible, as it treats detectors and classifiers as black boxes and learns from execution traces using reinforcement learning.
\end{abstract}

@@

\section{Conclusion}
We presented a method for learning ``closed-loop'' policies for multi-class object recognition, given existing object detectors and classifiers and a metric to optimize.
The method learns the optimal policy using reinforcement learning, by observing execution traces in training.
If detection on an image is cut off after only half the detectors have been run, our method does $66\%$ better than a random ordering, and $14\%$ better than an intelligent baseline.
In particular, our method learns to take action with no intermediate reward in order to improve the overall performance of the system.

As always with reinforcement learning problems, defining the reward function requires some manual work.
Here, we derive it for the novel detection AP vs. Time evaluation that we suggest is useful for evaluating efficiency in recognition.
Although computation devoted to scheduling actions is less significant than the computation due to running the actions, an interesting future research direction is to explicitly consider this decision-making cost; the same goes for feature computation costs.
The code for our method is available\footnote{\url{http://sergeykarayev.com/work/timely/}}.

\subsubsection*{Acknowledgments}
This research was made with Government support under and awarded by DoD, Air Force Office of Scientific Research, National Defense Science and Engineering Graduate (NDSEG) Fellowship, 32 CFR 168a.

### 2012-11-03 writeups/nips12/related.tex
camera-ready draft sent to collaborators

\subsection{Related Work}
To situate our contribution, let us briefly review approaches to object detection.

\myparagraph{Object detection}
The best recent performance has come from detectors that use gradient-based features to represent objects as either a collection of local patches or as object-sized windows \cite{Dalal2005,Lowe2004}.
Classifiers are then used to distinguish between featurizations of a given class and all other possible contents of an image window.
Window proposal is most often done exhaustively over the image space, as a ``sliding window''.
For state-of-the-art performance, the object-sized window models are augmented with parts \cite{Felzenszwalb2010a}, and the bag-of-visual-words models employ non-linear classifiers \cite{Vedaldi2009}.
%Some approaches use ``jump windows'' (hypotheses voted on by local features) \cite{Vedaldi2009,Vijayanarasimhan2011}, or a bounded search over the space of all possible windows \cite{Lampert2008a}.

@@

Most methods train individual models for each class.
Work on inherently multi-class detection focuses largely on making detection time sublinear in the number of classes through sharing features \cite{Torralba2007,Fan2005}.

@@

\myparagraph{Using context}
Leveraging \emph{context} for detection has a long-studied role in vision.
The most common source of context is the \emph{scene} or other non-detector cues; the most common scene-level feature is the GIST \cite{Oliva2001a} of the image.
For the commonly used PASCAL VOC dataset \cite{pascal-voc-2010}, GIST and other sources of context are quantitatively explored in~\cite{Divvala2009}.

@@

A critical summary of the main approaches to using context for object and scene recognition is given in \cite{Galleguillos2010}.

@@

To our knowledge, cascades have not been applied to multi-class detection, \todo{is this true?} and they are not dynamic policies.
They cannot change the order of execution based on observations obtained during execution, which is our goal.

>>>>

\subsection{Related Work}

\myparagraph{Object detection}
The best recent performance has come from detectors that use gradient-based features to represent objects as either a collection of local patches or as object-sized windows \cite{Dalal2005,Lowe2004}.
Classifiers are then used to distinguish between featurizations of a given class and all other possible contents of an image window.
Window proposal is most often done exhaustively over the image space, as a ``sliding window''.

For state-of-the-art performance, the object-sized window models are augmented with parts \cite{Felzenszwalb2010a}, and the bag-of-visual-words models employ non-linear classifiers \cite{Vedaldi2009}.
We employ the former detector in our evaluation.
%Some approaches use ``jump windows'' (hypotheses voted on by local features) \cite{Vedaldi2009,Vijayanarasimhan2011}, or a bounded search over the space of all possible windows \cite{Lampert2008a}.

@@

% Most detection methods train individual models for each class.
% Work on inherently multi-class detection focuses largely on making detection time sublinear in the number of classes through sharing features \cite{Torralba2007,Fan2005}.

@@

\myparagraph{Using context}
The most common source of context for detection is the \emph{scene} or other non-detector cues; the most common scene-level feature is the GIST \cite{Oliva2001a} of the image.
We use this source of scene context in our evaluation.

@@

A critical summary of the main approaches to using context for object and scene recognition is given in \cite{Galleguillos2010}.
For the commonly used PASCAL VOC dataset \cite{pascal-voc-2010}, GIST and other sources of context are quantitatively explored in~\cite{Divvala2009}.

@@

To our knowledge, cascades have not been applied to multi-class detection, and they are not dynamic policies.
They cannot change the order of execution based on observations obtained during execution, which is our goal.
